---
title: POL 304
subtitle: Introduction
author: "Olga Chyzh [www.olgachyzh.com]"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
exclude: true
editor_options: 
  chunk_output_type: console
---

```{r  setup, message=FALSE, warning=FALSE, include=FALSE}
options(
  htmltools.dir.version = FALSE, # for blogdown
  width = 80,
  tibble.width = 80
)

knitr::opts_chunk$set(
  fig.align = "center",  warning=FALSE, message=FALSE
)

```

```{r, echo=F}
library(tidyverse)
library(magrittr)

```


## Rules of Engagement

- Silence your mic unless you are speaking

- Raise your hand if you have a question/comment

- Keep your camera on if possible

- If you plan to follow asynchronously, email me *this week* to avoid participation penalties

---
## About me

- Assistant Professor of Political Science, since Fall 2020

- Previously an Assistant/Associate Professor of Political Science and Statistics at Iowa State University (2015-2020).

- My research areas are network analysis, spatial statistics, substantive focus on non-state violent actors

- [www.olgachyzh.com], olga.chyzh\@utoronto.ca

---
## Outline

- Objectives, software

- Assessments

- Getting started


---
## "Tall-Order" Objective

- Get introduced to a range of cutting-edge techniques for analyzing political and social data


---
## Objectives

- Learn the conceptual basics behind each technique

- Understand its applications

- Apply to real data

- Breadth over depth

---
## Set-up

- About 2:1 split of class between teaching:practice

- Recommend two screens: one to watch the presentation, one to follow along in R

---
## Assessments

- Participation, discussion of assigned readings --- 15\%

- 3 mini-projects (15\% each) and a final project (40\%)

- Apply the tools learned in class to your own data.

- As we go through applications, start thinking of your own examples

---
## Lectures

- Slides and all materials will be posted on Quercus at the beginning of each class

- Videos will be posted a couple days after class (time for processing)

- Follow along with in-class demos or you will get swamped

---
## What You Already Know

- Have you taken an Intro to Statistics? 

- Have you used R before?
 
- For more than a year?




